### 2011-01-28 trunk/papers/webist2011/paper_sec_catalogue.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Publishing and discovery platform}
\label{sec:discovery}

The area of web service publication and discovery has been subject for a lot of research since the very beginning the concept was coined. The current state of the art provides several solutions and strategies which providers and consumers may take advantage of. However, as explained in previous sections, a large number of them suffer from limited syntax-based descriptions and simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by non-technical end-users. In order to improve this issue, this paper presents a novel publishing platform, permitting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving web service wrappers for easy consumption in web applications.

\subsection{Overview}
\label{ssec:overview}

The main difference of this platform with regards to the state of the art is that it is targeting a different kind of user, and covers many deficiencies others tools do not tackle. As a brief overview, the platform being presented:

@@

This platform was intended and it is being used as part of the storyboard tool of the FAST platform~\cite{hoyer2009fast}, and it communicates with other components through its RESTful API, using JSON for the input of the requests. This paper will focus on the functionality regarding web service publication and discovery, and the process to publish any third-party web service into the platform. This process requires two steps:
\begin{inparaenum}[(i)]
	\item generation of the wrapper (see Sect. \ref{sec:wrapping_web_services}), and 
	\item creation of the resource in the publishing platform. 
\end{inparaenum}
The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been omitted, but the main idea is shown):

\begin{listing}
\begin{verbatim}
{
  "code": "http://demo.fast.morfeo-project.org/code/amazonSearch.js",
  "creationDate": "2010-01-26T17:01:13+0000",
  "description": {"en-gb": "Amazon web service - Search"},
  "label": {"en-gb": "Amazon search"},
  "actions": [{
    "name": "filter",
    "preconditions": [[{
      "id": "item",
      "label": {"en-gb": "Ebay List"},
      "pattern": "?item
                  http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                  http://aws.amazon.com/AWSECommerceService#Item",
      "positive": true
  }]],
  "postconditions": [
    ...
  ],
  ...
\end{verbatim}
\label{lis:json_request}
\end{listing}

The enriched search capabilities are supported by the definition of pre- and post-conditions. They allows to define the inputs and outputs of the services and other building blocks using concepts from any ontology, and in this way to find web services or other building blocks which match each other and can therefore be integrated. The concepts of pre-/post-conditions were strongly influenced by WSMO~\cite{roman2005}, simplified and implemented in RDFS for better live performance. The model used to define the web service wrappers and the ontology created are described in more detail in Sect.~\ref{ssec:ontology}.

\subsection{Platform architecture}
\label{ssec:architecture}

The architecture comprises an RDF store used for persistence, a business layer dealing with the model and reasoning, and a public facade providing the core functionality as a RESTful API, as well as a SPARQL endpoint accessing the RDF store directly. This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard (see Sect.~\ref{sec:use_case}), or any other third-party application.

The RESTful API and the SPARQL endpoint are part of the presentation layer. The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification~\cite{sprot} and is aimed to enable third-party developers to directly query the knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.

The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model specified in \ref{ssec:ontology},
acting as a mediator between the public RESTful API and the persistence layer.

The persistence layer provides an API allowing to work with a standard set of objects representing the model. The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against RDF2Go interfaces and choose any RDF store implementation. This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages.

\input{paper_ssec_ontology}

\input{paper_ssec_linked_data}

>>>>

\section{Publishing and Discovery Platform}
\label{sec:discovery}

As explained in previous sections, current solutions and strategies for web service publication and discovery suffer from 
limited syntax-based descriptions and simple keyword-based search, while other more complex approach failed because the added complexity did not offer sufficient benefits. 
This paper therefore presents a novel publishing platform (the \emph{catalogue}), permitting any enterprise or individual to publish their public web services, providing enhanced semantic search service wrappers for easy consumption in web applications.

% Overview subsection
\subsection{Overview}
\label{ssec:overview}

One of the main difference of this platform with regards to the state of the art is that it is targeting a different kind of user. As a brief overview, the platform being presented:

@@

% This paper will focus on the functionality regarding web service publication and discovery, and the process to publish any third-party web service into the platform. This process requires two steps:
% \begin{inparaenum}[(i)]
% 	\item generation of the wrapper (see Section \ref{sec:wrapping_web_services}), and 
% 	\item creation of the resource in the publishing platform. 
% \end{inparaenum}
% The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been omitted, but the main idea is shown):
% 
% \begin{listing}
% \begin{verbatim}
% {
%   "code": "http://fast.morfeo-project.org/amazon-search.js",
%   "creationDate": "2010-01-26T17:01:13+0000",
%   "description": {"en-gb": "Amazon web service - Search"},
%   "label": {"en-gb": "Amazon search"},
%   "actions": [{
%     "name": "filter",
%     "preconditions": [[{
%       "id": "item",
%       "label": {"en-gb": "Ebay List"},
%       "pattern": "?item
%         http://www.w3.org/1999/02/22-rdf-syntax-ns#type
%         http://aws.amazon.com/AWSECommerceService#Item"
%   }]],
%   "postconditions": [
%     ...
%   ],
%   ...
% \end{verbatim}
% \label{lis:json_request}
% \end{listing}

The enriched search capabilities are supported by the definition of pre- and post-conditions. They allow to define the inputs and outputs of the services and other building blocks using concepts from any ontology, and in this way to find web services or other building blocks which can be integrated. The concepts of pre-/post-conditions were strongly influenced by WSMO~\cite{roman2005}, simplified and implemented in RDFS for better live performance.

% % Architecture subsection
% \subsection{Platform architecture}
% \label{ssec:architecture}

The catalogue architecture comprises an RDF store used for persistence, a business layer dealing with the model and reasoning, and a public facade providing the core functionality as a RESTful API, as well as a SPARQL endpoint accessing the RDF store directly. This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard (see Fig.~\ref{fig:fast_architecture}), or any other third-party application.

% The RESTful API and the SPARQL endpoint are part of the presentation layer. The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification~\cite{sprot} and is aimed to enable third-party developers to directly query the knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.
% 
% The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model specified in \ref{ssec:ontology},
% acting as a mediator between the public RESTful API and the persistence layer.
% 
% The persistence layer provides an API allowing to work with a standard set of objects representing the model. The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against RDF2Go interfaces and choose any RDF store implementation. This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages.

% Conceptual model subsection
\input{paper_ssec_ontology}

% Discovery and recommendations subsection
\input{paper_ssec_discovery}

% Conceptual model subsection
\input{paper_ssec_linked_data}

### 2011-01-28 trunk/papers/webist2011/paper_sec_introduction.tex
%!TEX root = paper.tex
\subsection{Conceptual Model}
\label{ssec:ontology}
The conceptual model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and semantic approaches such as WSMO and OWL-S. It belongs to a more complex conceptual model for FAST, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom. 
In short, a screenflow is an executable (in a Web browser) set of screens, while a scree

Business-to-business (B2B) integration is still a significant challenge, often requiring extensive efforts in terms of different aspects and technologies for protocols, architectures or security. Solutions based on open standards have reduced the complexity of integrating different business applications between different companies and partners. 
In this sense, Web services and XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B. By web services standards we are referring to the following open standards: Web Services Description Language (WSDL --- to describe), Universal Description, Discovery and Integration (UDDI --- to advertise and syndicate), Simple Object Access Protocol (SOAP --- to communicate) and Web Services Flow Language (WSFL --- to define work flows). 
A common scenario  in their usage is a company using WSDL to describe its web services, publishing them through a repository using UDDI, while these web services use SOAP-based messages to achieve dynamic integration between different, disparate applications.

While the adoption of these standards has offered some advantages in business-to-consumer (B2C) integration as well,  the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge on the matter (in the sense as defined in~\cite{fuchsloch2010}). 
Regarding the discovery of the right service for the right task --- a crucial requirement for the dynamic use of WS --- most publishing platforms are syntax-based, %leading to poor precision and recall in finding the correct service, 
making it difficult to navigate through a large number of web services~\cite{pilioura_acm2009}, 
therefore preventing end-users from performing these tasks. 
Solutions such as Semantic Web Services (SWS) promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, location, service implementation or data format. 
However, as of yet they have not been widely adopted, 
possibly because the perceived potential benefits of SWS did not justify the additional investments required for businesses~\cite{shi2007}.
In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
So far, the efforts are restricted to data only, so that users are left alone in finding applications of that data. 
A natural further development in the public sector would be that their services for administration tasks and single-window applications will be opened as well. Hence, governments will benefit from the research done around these concepts and technologies.

The motivation of our work has been strongly influenced by the end-users' needs. We are targeting users which are non-skilled in term of programming and software development, empowering them with a platform to discover and consume web services in a straight-forward manner.
The web services and their specifications are not published right away in the platform, since this would simply mean to reimplement UDDI.
However, their definition and behaviour are wrapped in what we call a ``web service wrapper''. The products resulting from the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-consume (inside a web browser) piece of code with the proper functions to invoke the web service.

The remainder of this paper is organised as follows. Section~\ref{sec:related_work} covers the state of the art in web services publication, discovery and wrapping/consumption. In Section~\ref{sec:wrapping_web_services} we present our approach for web services wrapping, available possibilities and limitations. Next, Sect.~\ref{sec:discovery} describes our platform for publishing web services and empowering end-users to discover services. Section~\ref{sec:use_case} gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section~\ref{sec:conclusions}.

>>>>

Business-to-business (B2B) integration is still a significant challenge, often requiring extensive efforts in terms of different aspects and technologies for protocols, architectures or security. Solutions based on open standards 
such as RosettaNet, ebXML, the Web Service Description Language (WSDL), Universal Description, Discovery and Integration (UDDI) or the Simple Object Access Protocol (SOAP)
have reduced the complexity of integrating different business applications between different companies and partners. 
% In this sense, Web services and XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B. By web services standards we are referring to the following open standards: Web Services Description Language (WSDL --- to describe), Universal Description, Discovery and Integration (UDDI --- to advertise and syndicate), Simple Object Access Protocol (SOAP --- to communicate) and Web Services Flow Language (WSFL --- to define work flows). 
% A common scenario  in their usage is a company using WSDL to describe its web services, publishing them through a repository using UDDI, while these web services use SOAP-based messages to achieve dynamic integration between different, disparate applications.

While the adoption of these standards has offered some advantages in business-to-consumer (B2C) integration as well,  the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge on the matter.% (in the sense as defined in~\cite{fuchsloch2010}). 
Regarding the discovery of the right service for the right task --- a crucial requirement for the dynamic use of WS --- most publishing platforms are syntax-based, %leading to poor precision and recall in finding the correct service, 
making it difficult to navigate through a large number of web services~\cite{pilioura_acm2009}, 
therefore preventing end-users from performing these tasks. 
Solutions such as Semantic Web Services (SWS) promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, location, service implementation or data format. 
However, as of yet they have not been widely adopted, 
possibly because the perceived potential benefits of SWS did not justify the additional investments required for businesses~\cite{shi2007}.
In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
So far, the efforts are restricted to data only, so that users are left alone in finding applications of that data. 
A natural further development in the public sector would be that their services for administration tasks and single-window applications will be opened as well. Hence, governments will benefit from the research done around these concepts and technologies.

The motivation of our work has been strongly influenced by the end-users' needs. We are targeting users which are non-skilled in term of programming and software development, empowering them with a platform to discover and consume web services in a straight-forward manner.
The web services and their specifications are not published right away in the platform, since this would simply mean to re-implement UDDI.
However, their definition and behaviour are wrapped in what we call a ``web service wrapper''. The products resulting from the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-consume (inside a web browser) piece of code with the proper functions to invoke the web service.

% The remainder of this paper is organised as follows. Section~\ref{sec:related_work} covers the state of the art in web services publication, discovery and wrapping/consumption. In Section~\ref{sec:wrapping_web_services} we present our approach for web services wrapping, available possibilities and limitations. Next, Section~\ref{sec:discovery} describes our platform for publishing web services and empowering end-users to discover services. Section~\ref{sec:use_case} gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section~\ref{sec:conclusions}.

### 2011-01-28 trunk/papers/webist2011/paper_sec_related_work.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Related work}
\label{sec:related_work}

@@

A different approach for web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification~\cite{beatty2005}. The core of this approach is a multicast discovery protocol. Service providers and consumers listen to each other for new services specifications within a network, so there is no need of a centralised registry. As a drawback, WS-Discovery do not support Internet-scale discovery, making it useless for our purposes.

>>>>

\section{Related Work}
\label{sec:related_work}

@@

% A different approach for web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification~\cite{beatty2005}. The core of this approach is a multicast discovery protocol. Service providers and consumers listen to each other for new services specifications within a network, so there is no need of a centralised registry. As a drawback, WS-Discovery do not support Internet-scale discovery, making it useless for our purposes.

### 2011-01-28 trunk/papers/webist2011/paper_sec_use_case.tex
adapted to WEBIST format
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@980 d8b0c59b-6544-0410-94d9-83a5658a058e
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Use case}
\label{sec:use_case}

The intention of this paper was to show the advances made regarding web service discovery and consumption through the two artifacts developed as part of our research. While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform.
FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective. Therefore, it is aimed at allowing users (mainly business users at the management level, without programming experience) to create their own situational applications by visually combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs. The work covered in this paper are the components highlighted and surronded by a dashed frame in the Fig.~\ref{fig:fast_architecture}. These components have public interfaces to communicate with any application developed by a third-party developer or company, and as described in Sects. 3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture.

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.9\linewidth]{images/fast_architecture.pdf}
    \caption{Overview of the FAST architecture}
    \label{fig:fast_architecture}
  \end{center}
\end{figure}

In order to facilitate a better understanding of the use case, we will describe a number of concepts related to FAST in this section. A \emph{gadget} is the end product of the platform, aimed to be used in a specific mashup platform. Within FAST, the platform-independent gadget is called \emph{screenflow}, which comprises a set of domain-specific \emph{screens} connected through their \emph{pre-} and \emph{post-conditions}. A screen is the most complex building block fully functional by itself. It is composed by a \emph{form} conveying the graphical interface, and a set of \emph{operators} and \emph{back-end services}, wrapped into so-called \emph{service resources}.

The web service wrapper tool is then used to create the formal definitions of back-end services, transforming them into building blocks to be stored and reused within the FAST platform. The tool is integrated directly into the graphical IDE, and for the storage interacts with the catalogue, where the formal service descriptions and the executable code encapsulating the service behaviour is stored.

The publishing platform presented, also called ``Catalogue'' within the FAST platform, covers several important purposes:
\begin{description}
	\item[Publication and search of gadgets] The output of the FAST IDE are gadgets, which are self-contained and ready to use. They need to be made available for download, as well as being discovered according to their formal description.
	\item[Publication and search of building blocks] Every single component or building block definition is stored in the catalogue, which facilitates searching, recommendation and planning (the suggestion of combinations of building blocks to fulfil certain requirements). 
	\item[Storage and indexing of user profiles and histories] Each user of the FAST platform has a representation in the catalogue, thus making it possible for the IDE to make informed guesses about what they want in a particular situation, based on their previous behaviour.
	\item[Facilitate ontology mediation] Because FAST allows the integration of third-party external services, it is not always guaranteed that all the data coming from a service is directly compatible with the FAST platform. In such cases, mediation between the different ontologies is necessary. The catalogue provide some grade of mediation functionality (which is not covered in this paper).
\end{description}

>>>>

\section{Context: The FAST Platform}
\label{sec:use_case}

The intention of this paper is to demonstrate the advances made regarding web service discovery and consumption through two artifacts developed as part of our research: the \emph{publishing and discovery platform} and the \emph{service wrapper tool}. While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform~\cite{hoyer2009fast}.
FAST constitutes a novel approach to application composition and business process definition from a user-centric perspective. It is aimed at allowing users without previous programming experience to create their own situational applications by visually combining different \emph{building blocks}, such as graphical forms and back-end services, based on their inputs and outputs (or \emph{pre-} and \emph{post-conditions}). The work covered in this paper are the components highlighted and surrounded by a dashed frame in the Fig.~\ref{fig:fast_architecture}. Communication within the platform is mostly done via a RESTful API, using JSON as an exchange syntax.
%These components have public interfaces to communicate with any application developed by a third-party developer or company, and as described in Sects. 3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture.

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=\linewidth]{images/fast_architecture.pdf}
    \caption{Overview of the FAST architecture}
    \label{fig:fast_architecture}
  \end{center}
\end{figure}

In order to facilitate a better understanding of these components, we describe a number of concepts related to FAST in this section. 
A \emph{gadget} is the end product of the platform, ready to be run in any ordinary web browser, usually through a mashup platform. 
Within FAST, the platform-independent gadget is called \emph{screenflow}, which comprises a set of domain-specific \emph{screens} connected through their \emph{pre-} and \emph{post-conditions}. A screen is the most complex building block fully functional by itself. 
Visually, it is similar to a tab in a tabbed application.
It is composed by a \emph{form} conveying the graphical interface, and a set of \emph{operators} and \emph{back-end services}, wrapped into so-called \emph{service resources}.

The service wrapper tool is used to create wrappers for third-party web services, transforming them into building blocks to be stored and reused within the FAST platform. The publishing platform presented, also called \emph{catalogue} within the FAST platform, covers several important purposes, such as storage, indexing, publication and search of gadgets, gadget building blocks and user profiles, as well as some level of ontology mediation to facilitate the integration of services from different parties (first steps in this direction are outlined in \cite{Ambrus:2009it}).

### 2011-01-28 trunk/papers/webist2011/paper_sec_wrapping.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Wrapping web services}
\label{sec:wrapping_web_services}

@@

\subsection{Constructing service requests} % (fold)
\label{sub:constructing_service_requests}

In this paper we illustrate the interaction with services on simple GET requests as supported e.g. by REST services. Support for other services like POST based services is under construction. In case of a simple GET request, a service request is assembled using a certain URL and a set of parameters. (A POST request would have a text block to be posted. In addition, other header data may be needed to transfer e.g. session ids.) As an example, the Ebay Shopping web service will be studied. As documented on the corresponding web site, the following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}:

@@

As you may see, the URL invoked is \url{http://open.api.sandbox.ebay.com/shopping} and the parameters used in the example are:
\begin{description}
	\item[appid] this is the application ID obtained to use the API.
	\item[version] the API version.
	\item[callname] in this case FindItems to search through all items in Ebay.
	\item[itemsort] sorting method for the list of items.
	\item[querykeywords] list of keywords.
	\item[responseencoding] format of the response message obtained by the invocation of the request.
\end{description}

@@

To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web based form that allows the editing of these entries, see Fig.~\ref{fig:construct_pre_post_conditions}. Note, we allow to edit example values for the input parameters. These example values may be used to test the service wrapper.

@@

The service wrapper tool composes the request using a template string which will contain placeholders for input values. Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent as the service request.

An example of service request construction can be found in the Figure~\ref{fig:construct_service_request}. In the top input field the user may drop an example HTTP request taken e.g. form the service documentation e.g. from its website\footnote{http://developer.ebay.com/products/shopping/}. The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided. In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter. In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.

@@

Below the request parameter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay). Then the placeholders for input parameters are replaced by their example values and the resulting http request is shown below the parameter form. In addition, the request is sent and the response is shown on the bottom of that page. This gives the user a fast feedback whether the constructed request works as desired.

@@

\subsection{Interpreting service responses} % (fold)
\label{sub:interpreting_service_responses}

The service request is constructed and sent to the service provider, who will then send back a response. This response message could be serialised in any format, though the most common formats used nowadays are XML or JSON. To continue the example started in the previous section, we assume the response of the Ebay Shopping service will be in XML as specified in the request.

\subsubsection{Translation of XML into Facts} % (fold)
\label{ssub:translation_xml_into_facts}

@@

Once the service response, in XML format, has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen on the left side of Fig.~\ref{fig:response_service_execution}. To construct this interactive object tree, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.

A transformation rule is used to analyse the XML data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks. A transformation rule is composed of three elements, as seen in the middle part of Fig.~\ref{fig:response_service_execution}. First, the \textit{from} field indicates the XML elements to be translated by the rule. These XML elements are identified by the tag name of a DOM element from the XML document. Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}. And third, the target of the rule specifies a certain concept or attribute, to be created or filled. A detailed explanation of the type of actions to be trigger from the transformation rules is as follows:
\begin{description}
	\item[action \emph{createObject}] specifies the creation of a new domain object. The type of that new object is provided in the third compartment. In the example being explained, the root rule searches for XML elements with tag name \emph{FindItemsResponse} and for each such element a \emph{List} object is created. The resulting objects are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
	\item[action \emph{fillAttributes}] does not create a new object but it fills the value of the attribute provided as third part of such rules. In our example, the third transformation rule searches for XML elements with tag name \emph{Title}. Note, the rule is a sub-rule of the second rule, which generates \emph{Product} objects. Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data. Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ). The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} object. Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g. to \textit{words} 1 through 3. It is also possible to combine constant text and elements of multiple XML tree elements. 
	\item[action \emph{dummy}] does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules. For example, in Amazon product data, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such section contains the \emph{plain price} and the \emph{formatted price}. Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
\end{description}

Our tool follows an interactive paradigm. Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, our tool is ontology-driven, therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Section~\ref{sec:discovery}) together with the structure of each type, i.e. together with a description of the attributes of each object. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the object types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule. In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Section~\ref{ssec:ontology}). This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

>>>>

\section{Wrapping Web Services}
\label{sec:wrapping_web_services}

@@

\subsection{Constructing Service Requests} % (fold)
\label{sub:constructing_service_requests}

In this paper we illustrate the interaction with services on simple GET requests as supported e.g. by REST services. Support for other services like POST based services is under construction. In case of a simple GET request, a service request is assembled using a certain URL and a set of parameters. (A POST request would have a text block to be posted. In addition, other header data may be needed to transfer e.g. session ids). As an example, the Ebay Shopping web service will be studied. As documented on the corresponding web site, the following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}:

@@

As you may see, the URL invoked is \url{http://open.api.sandbox.ebay.com/shopping} and the parameters used in the example are:
\begin{description}
	\item[appid] this is the application ID obtained to use the API.
	\item[version] the API version.
	\item[callname] in this case \emph{FindItems} to search through all items in Ebay.
	\item[ItemSort] sorting method for the list of items.
	\item[QueryKeywords] list of keywords.
	\item[responseencoding] format of the response message obtained by the invocation of the request.
\end{description}

@@

To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web-based form that allows the editing of these entries, see Fig.~\ref{fig:construct_pre_post_conditions}. Note, we allow to edit example values for the input parameters. These example values may be used to test the service wrapper.

@@

% The service wrapper tool composes the request using a template string which will contain placeholders for input values. Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent as the service request.

An example of service request construction can be found in the Figure~\ref{fig:construct_service_request}. In the top input field the user may drop an example HTTP request, taken e.g. from the service's documentation\footnote{\url{http://developer.ebay.com/products/shopping/}}. The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided. In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter. In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.

@@

Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay). 
The service's response is shown on the bottom of the page. This gives the user a fast feedback whether the constructed request works as desired.

@@

\subsection{Interpreting Service Responses} % (fold)
\label{sub:interpreting_service_responses}

A web service can choose to send a response message in any desired structured format. In our example, we assume an XML tree, but a JSON structure is equally possible.

\subsubsection{Translation of the Response Tree into Facts} % (fold)
\label{ssub:translation_xml_into_facts}

@@

Once the service response has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen on the left side of Fig.~\ref{fig:response_service_execution}.

A transformation rule is used to analyse the response data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks. A transformation rule is composed of three elements, as seen in the middle part of Fig.~\ref{fig:response_service_execution}. First, the \textit{from} field indicates the source elements to be translated by the rule. Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}. And third, the target of the rule specifies a certain concept or attribute, to be created or filled. A detailed explanation of the type of actions to be trigger from the transformation rules is as follows:

\paragraph{createObject} % (fold)
\label{par:createobject}

specifies the creation of a new domain object. The type of that new object is provided in the third compartment. In the example being explained, the root rule searches for XML elements with tag name \emph{FindItemsResponse} and for each such element a \emph{List} object is created. The resulting objects are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.

% paragraph createobject (end)

\paragraph{fillAttributes} % (fold)
\label{par:fillattributes}

does not create a new object but it fills the value of the attribute provided as third part of such rules. In our example, the third transformation rule searches for XML elements with tag name \emph{Title}. Note, the rule is a sub-rule of the second rule, which generates \emph{Product} objects. Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data. Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ). The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} object.

% paragraph fillattributes (end)

\paragraph{dummy} % (fold)
\label{par:dummy}

does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules. For example, in Amazon product data, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such section contains the \emph{plain price} and the \emph{formatted price}. Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.

% paragraph dummy (end)

Our tool follows an interactive paradigm. Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, our tool is ontology-driven, therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Section~\ref{sec:discovery}) together with the structure of each type, i.e. together with a description of the attributes of each object. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the object types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule. 
% In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Section~\ref{ssec:ontology}). This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.
